initial_analysis_notebook

### 2021-04-27 exploratory/notebook/initial_analysis/initial_analysis_notebook.Rmd
add lefse and interactive feature plot

---
title: "initial_analysis_notebook"
author: "Nick Lesniak"
date: "4/13/2021"
output: rmarkdown::github_document
---

@@

```{r}
metadata <- read_tsv('../../../data/process/metadata_tidy.tsv',
		col_type = 'cccccDddddDcdddddcdl') 
shared <- read_tsv('../../../data/mothur/sample.final.0.03.subsample.shared',
		col_types = cols(.default = col_double(),
			Group = col_character())) %>% 
	mutate(Group = ifelse(grepl('_D|DA', Group), Group,
		gsub('D', '_D', Group)))
taxonomy <- read_tsv('../../../data/process/final.taxonomy.tidy.tsv',
                     col_types = cols(.default = col_character()))

@@

lefse_plot <- lefse_df %>% 
	ggplot(aes(y = LDA, fill = Class, 
			x = reorder(tax_otu_label, order))) +
		geom_col() + 
		theme_bw() + 
		coord_flip() + 
		guides(fill = 'none') + 
		scale_fill_manual(values = c(1,2),
			limits = c(TRUE, FALSE),
			labels = c('Severe', 'Non-severe')) + 
		labs(x = NULL, y = 'LDA Score (log 10)')

abundance_plot <- metadata %>% 
	filter(cdiff_strain == 431,
		day == 0) %>% 
	left_join(shared, by = c('group' = 'Group')) %>% 
	gather(OTU, abundance, matches('Otu\\d*')) %>% 
	left_join(taxonomy, by = 'OTU') %>% 
	right_join(lefse_df, by = c('Genus' = 'OTU')) %>% 
	group_by(group, Genus, order, early_euth) %>% 
	summarise(abundance = sum(abundance)) %>% 
	mutate(tax_otu_label = Genus) %>% 
	mutate(abundance = abundance / 21.07,
		abundance = ifelse(abundance == 0, 0.045, abundance)) %>% 
	ggplot(aes(x = reorder(tax_otu_label, order), 
			y = abundance, color = as.factor(early_euth))) + 
		geom_jitter(position = position_jitterdodge(dodge.width = 0.5), alpha = 0.25) + 
		stat_summary(fun.data = 'median_hilow', aes(group = early_euth),
			fun.args = (conf.int=0.5), position = position_dodge(width = 0.5)) +
		theme_bw() + labs(x = NULL, y = 'Relative Abundance', color = NULL) + 
		scale_y_log10(breaks=c(0.01, 0.1, 1, 10, 100),
			labels=c("0","0.1","1","10","100")) +
		scale_color_manual(values = c(1,2),
						   limits = c(TRUE, FALSE),
						   labels = c('Severe', 'Non-severe')) + 
		geom_hline(yintercept = 0.047, linetype = 'dashed', lwd = 0.3) + 
		theme(panel.grid.major.y = element_blank(), 
			panel.grid.minor = element_blank(),
			axis.ticks.y = element_blank(),
			axis.text.y = element_blank(),
			legend.position = 'right') + 
		coord_flip()

cowplot::plot_grid(lefse_plot, abundance_plot, nrow = 1)

>>>>

---
title: "initial_analysis_notebook"
author: "Nick Lesniak"
date: "4/13/2021"
always_allow_html: true
output: 
  html_document: default
---

@@

```{r}
library(plotly)
metadata <- read_tsv('../../../data/process/metadata_tidy.tsv',
		col_type = 'cccccDddddDcdddddcdl') 
shared <- read_tsv('../../../data/mothur/sample.final.0.03.subsample.shared',
		col_types = cols(.default = col_double(),
			Group = col_character())) %>% 
	mutate(Group = ifelse(grepl('_D|DA', Group), Group,
		gsub('D', '_D', Group)))
taxonomy <- read_tsv('../../../data/process/final.taxonomy.tidy.tsv',
                     col_types = cols(.default = col_character()))

@@

lefse_plot <- lefse_df %>% 
	ggplot(aes(y = LDA, fill = Class, 
			x = reorder(tax_otu_label, order))) +
		geom_col() + 
		theme_bw() + 
		coord_flip() + 
		guides(fill = 'none') + 
		scale_fill_manual(values = c('#8de4d3','#f75b5f'),
			limits = c(TRUE, FALSE),
			labels = c('Severe', 'Non-severe')) + 
		labs(x = NULL, y = 'LDA Score (log 10)')

abundance_plot <- metadata %>% 
	filter(cdiff_strain == 431,
		day == 0) %>% 
	left_join(shared, by = c('group' = 'Group')) %>% 
	gather(OTU, abundance, matches('Otu\\d*')) %>% 
	left_join(taxonomy, by = 'OTU') %>% 
	right_join(lefse_df, by = c('Genus' = 'OTU')) %>% 
	group_by(mouse_id, group, Genus, order, early_euth) %>% 
	summarise(abundance = sum(abundance)) %>% 
	mutate(tax_otu_label = Genus) %>% 
	mutate(abundance = abundance / 21.07,
		abundance = ifelse(abundance == 0, 0.045, abundance)) %>% 
  highlight_key(~mouse_id) %>% 
	ggplot(aes(x = reorder(tax_otu_label, order), 
			y = abundance, color = as.factor(early_euth), shape = mouse_id)) + 
		stat_summary(fun.data = 'median_hilow', aes(group = early_euth),
			fun.args = (conf.int=0.5), position = position_dodge(width = 0.5)) +
		theme_bw() + labs(x = NULL, y = 'Relative Abundance', color = NULL) + 
		scale_y_log10(breaks=c(0.01, 0.1, 1, 10, 100),
			labels=c("0","0.1","1","10","100")) +
		scale_color_manual(values = c('#8de4d3','#f75b5f'),
						   limits = c(TRUE, FALSE),
						   labels = c('Severe', 'Non-severe')) + 
    scale_shape_manual(values = rep(16, 54)) + 
		geom_hline(yintercept = 0.047, linetype = 'dashed', lwd = 0.3) + 
		theme(panel.grid.major.y = element_blank(), 
			panel.grid.minor = element_blank(),
			legend.position = 'right') + 
		coord_flip() + 
    guides(shape = 'none')

cowplot::plot_grid(lefse_plot, 
                   abundance_plot + 
                     geom_jitter(position = position_jitterdodge(dodge.width = 0.5), alpha = 0.25) + 
                     theme(axis.ticks.y = element_blank(),
                           axis.text.y = element_blank()), 
                   nrow = 1)

gg <- ggplotly(abundance_plot + geom_jitter(position = position_jitterdodge(dodge.width = 0.5)), tooltip = c('mouse_id'))
highlight(hide_legend(gg), on = "plotly_click", off = "plotly_doubleclick", 
           color = "black", selectize = T, opacityDim = .2,
           selected = attrs_selected(showlegend = F, marker = list(symbol = 'X')))
str(gg)
```
